---
title: "Example Parameter Simulations"
author: "Kassel Hingee"
date: "16/03/2022"
output: 
  html_document: 
    toc: yes
---

This document contains plots of a number of example PPI models.

```{r setup, include=FALSE}
library(scorecompdir)
library(Ternary)
library(ggtern)
library(ggplot2)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

# Model from Section 2.3 of JASA Paper
```{r simulateS2_3model}
#dimension
p=3

#sample size
n=1000

#parameters for the PPI model
muL=matrix(0,p-1,1)
muL[1:sum(p,-1)]=0.12 #location is close to u1=0, u2=0.
aa=matrix(1/500,p-1,1)
D=diag(as.vector(aa))
SigA=D
SigA[1,1]=SigA[1,1]*2
cor=0.5
SigA[1,2]=cor*sqrt(SigA[1,1]*SigA[2,2])
SigA[2,1]=SigA[1,2]
ALs=-0.5*solve(SigA)
bL=solve(SigA)%*%muL
beta0=matrix(-0.8,p,1)
beta0[p]=-0.5

#simulate sample from PPI model
samp=scorecompdir:::rppi_singly(n,p,beta0,ALs,bL,4)[[1]]
# samp=rppi(n,beta = beta0, AL = ALs, bL = bL,maxden = 4)
```

```{r plot1_ternarypackage}
colMeans(samp)
Ternary::TernaryPlot("u1", "u2", "u3",
                     "larger u1 \U2192")
# Ternary::ColourTernary(Ternary::TernaryDensity(coordinates = samp$samp3) )
Ternary::AddToTernary(points, samp, pch = "+")
```

```{r s2_3modelplot}
data.frame(samp) %>%
  dplyr::rename(u1 = X1, u2 = X2, u3 = X3) %>%
  ggplot(aes(x = u1, y = u2, z = u3)) +
  coord_tern() +
  geom_density_tern(bins = 5, bdl = 0.01) + 
  geom_point(size = 1, alpha = 0.5, shape = "+")
```

# Centre of Simplex
For a model with points focused on the interior of the simplex, change muL (influences bL) to close to 0.5, and make the beta exponents large.

```{r centresimplex_simulate}
#parameters for the PPI model
muL=matrix(0,p-1,1)
muL[1:sum(p,-1)]=0.3
aa=matrix(1/500,p-1,1)
D=diag(as.vector(aa))
SigA=D
SigA[1,1]=SigA[1,1]*2
cor=0.5
SigA[1,2]=cor*sqrt(SigA[1,1]*SigA[2,2])
SigA[2,1]=SigA[1,2]
ALs=-0.5*solve(SigA)
bL=solve(SigA)%*%muL
beta0=matrix(1,p,1)
samp=scorecompdir:::rppi_singly(n,p,beta0,ALs,bL,4)[[1]]
# samp=rppi(n,beta = beta0, AL = ALs, bL = bL,maxden = 4)
```

```{r centresimplex_plot}
colMeans(samp)
data.frame(samp) %>%
  dplyr::rename(u1 = X1, u2 = X2, u3 = X3) %>%
  ggplot(aes(x = u1, y = u2, z = u3)) +
  coord_tern() +
  geom_density_tern(bins = 5, bdl = 0.01) + 
  geom_point(size = 1, alpha = 0.5, shape = "+")
```

# Centre of Simplex with Near Zeros
An implausible model with points that are interior to the simplex and extra density near the simplex edges.
```{r centresimplex_wedge_simulate}
#parameters for the PPI model
muL=matrix(0,p-1,1)
muL[1:sum(p,-1)]=0.3
aa=matrix(1/500,p-1,1)
D=diag(as.vector(aa))
SigA=D
SigA[1,1]=SigA[1,1]*2
cor=0.5
SigA[1,2]=cor*sqrt(SigA[1,1]*SigA[2,2])
SigA[2,1]=SigA[1,2]
ALs=-0.5*solve(SigA)
bL=solve(SigA)%*%muL
beta0=matrix(-0.8,p,1)
samp=scorecompdir:::rppi_singly(n,p,beta0,ALs,bL,4)[[1]]
# samp=rppi(n,beta = beta0, AL = ALs, bL = bL,maxden = 4)
```

```{r centresimplex_w_edge_plot}
colMeans(samp)
data.frame(samp) %>%
  dplyr::rename(u1 = X1, u2 = X2, u3 = X3) %>%
  ggplot(aes(x = u1, y = u2, z = u3)) +
  coord_tern() +
  geom_density_tern(bins = 5, bdl = 0.01) + 
  geom_point(size = 1, alpha = 0.5, shape = "+")
```

# Large u1
```{r largeu1}
#parameters for the PPI model
muL=matrix(0,p-1,1)
muL[1, 1] <- 0.9
muL[2, 1] <- 0.12
aa=matrix(1/500,p-1,1)
D=diag(as.vector(aa))
SigA=D
SigA[1,1]=SigA[1,1]*2
cor=0.5
SigA[1,2]=cor*sqrt(SigA[1,1]*SigA[2,2])
SigA[2,1]=SigA[1,2]
ALs=-0.5*solve(SigA)
bL=solve(SigA)%*%muL
beta0=matrix(c(-0.5, -0.5, -0.5),p,1)

#simulate sample from PPI model
samp=scorecompdir:::rppi_singly(n,p,beta0,ALs,bL,4)[[1]]
# samp=rppi(n,beta = beta0, AL = ALs, bL = bL,maxden = 4)
```

```{r largeu1_plot}
data.frame(samp) %>%
  dplyr::rename(u1 = X1, u2 = X2, u3 = X3) %>%
  ggplot(aes(x = u1, y = u2, z = u3)) +
  coord_tern() +
  geom_density_tern(bins = 5, bdl = 0.01) + 
  geom_point(size = 1, alpha = 0.5, shape = "+")
```

# Dirichlet
```{r dirichlet_simulate}
beta0=matrix(-0.8,p,1)
beta0[p]=-0.5

#simulate sample from PPI model
samp=scorecompdir:::rppi_singly(n,p,beta0,matrix(0, 2, 2), matrix(0, 2, 1),4)[[1]]
# samp=rppi(n,beta = beta0, AL = 0, bL = 0,maxden = 4)
```

```{r dirichlet_plot}
data.frame(samp) %>%
  dplyr::rename(u1 = X1, u2 = X2, u3 = X3) %>%
  ggplot(aes(x = u1, y = u2, z = u3)) +
  coord_tern() +
  geom_density_tern(bins = 5, bdl = 0.01) + 
  geom_point(size = 1, alpha = 0.5, shape = "+")
```
